Modern BTree techniques

### 2022-09-17 strangeloop/presentation.tex
Rearrange table
Adjustments

    \begin{itemize}[label={\MVRightarrow}]
        \item <+-> D. Comet, "Ubiquitous B-tree", ACM Comp. Surv., vol. 11, no.
            2, 1979.
        \item <+-> Graefe, Goetz and Harumi A. Kuno. "Modern B-tree
            techniques." IEEE 27th International Conference on Data
            Engineering, 2011.
    \end{itemize}
\end{frame}

@@

    \begin{itemize}[label={\MVRightarrow}]
        \item RUM conjecture
        \item B-Tree
        \begin{itemize}[label={}]
            \item Key normalization \& friends
            \item SB-Tree
            \item MDAM
        \end{itemize}
        \item Hybrid, Bw-Tree, DPTree
        \item Trie
        \item Everything else
    \end{itemize}
\end{frame}

    \def\arraystretch{1.5}
    \begin{tabular}{cccc}
        B-Tree & B$^{+}$-Tree & B$_{link}$-Tree & DPTree \\
        wB$^{+}$-Tree & NV-Tree & FPTree & FASTFAIR \\
        HiKV & Masstree & Skip List & ART \\
        WORT & CDDS-Tree & Bw-Tree & HOT \\
        KISS-Tree & VAST-Tree & FAST & HV-Tree \\
        UB-Tree & LHAM & MDAM & Hybrid B$^{+}$-Tree
    \end{tabular}
\end{frame}

@@

@@

    \begin{center}
    \textbf{B$^+$Tree}
    \vspace{1cm}

@@

            \draw[->, btree-line] ([yshift=0.1cm]sep33.east) -- ([yshift=0.1cm]sep42.west);
            \draw[->, btree-line] ([yshift=0.1cm]sep32.east) -- ([yshift=0.1cm]sep44.west);
            \draw[<-, btree-line] ([yshift=-0.1cm]sep33.east) -- ([yshift=-0.1cm]sep42.west);
            \draw[<-, btree-line] ([yshift=-0.1cm]sep32.east) -- ([yshift=-0.1cm]sep44.west);
            \coordinate[right=0.5cm of sep34.east] (link-sep34);
            \draw[->, btree-line] ([yshift=0.1cm]sep34.east) -- ([yshift=0.1cm]link-sep34.west);
            \coordinate[left=0.5cm of sep43.west] (link-sep43);
            \draw[->, btree-line] ([yshift=0.1cm]sep43.west) -- ([yshift=0.1cm]link-sep43.east);
        \end{tikzpicture}
    \end{overprint}
    \end{center}

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        P. Lehman and S. Yao, Efficient Locking for Concurrent Operations on
        B-Trees, ACM Transactions on Database Systems, Vol 6, No. 4, December
        1981, pp 650-670
    \linespread{1.5}

@@

    \begin{flushleft}
        \begin{displayquote}
            "By adding periodic rebuilding of the tree, we obtain a data
            structure that is theoreticaly superior to standard B-trees in many
            ways. Our results suggest that rebalancing on deletion no only
            unnecessary but may be harmful."
        \end{displayquote}
        \footnotesize{S. Sen and R. E. Tarjan, "Deletion without rebalancing in multiway
        search trees." ISAAC, pp. 832-841, 2009}
    \end{flushleft}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        R. C. Singleton, "An efficient algorithm for sorting with minimal
        storage", Communications of the ACM, vol. 12, no.3 pp. 185-186, 1969
    \linespread{1.5}

@@

        \onslide<5>
        \begin{center}
        \begin{tikzpicture}[]
            \btreenodewithitems{1}{8}{5}{}{btree-leaf}
            \foreach \i in {1,...,4}
            {
                \coordinate[below=1cm of key1\i] (start\i);
                \draw[->, btree-line, line width=1.5pt, shorten >= 2mm] (start\i) -- (key1\i);
            }
        \end{tikzpicture}
        \end{center}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        R. Bayer, K. Unterauer, "Prefix B-trees", ACM Transactions on Database
        Systems, vol. 2, no. 1, pp. 11-26, 1977
    \linespread{1.5}

    \end{center}
\end{frame}
\note{
    Not a normalized version, but whole column version is in PG12
}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        G. Graefe, Per-Ake Larson, "B-tree indexes and CPU caches",
        International Conference on Data Engineering, pp. 349-358, 2001
    \linespread{1.5}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        P. E. O'Neil, "The SB-tree: An index-sequential structure for high
        performance sequential access", Acta Informatica, vol. 29, no. 3. pp.
        241-265, 1992
    \linespread{1.5}

@@

    \begin{center}
    \textbf{Multidimentional Access Method}
    \vspace{1cm}

    \begin{overprint}[8cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \basicbtree{show-connections}
            \node[below=0.2 of node5] {\footnotesize{1,a}};
            \node[below=1.8 of node2] {\footnotesize{...}};
            \node[below=0.2 of node6] {\footnotesize{1,z}};

            \node[below=0.2 of node7] {\footnotesize{2,a}};
            \node[below=1.8 of node3] {\footnotesize{...}};
            \node[below=0.2 of node8] {\footnotesize{2,z}};

            \node[below=0.2 of node9] {\footnotesize{3,a}};
            \node[below=1.8 of node4] {\footnotesize{...}};
            \node[below=0.2 of node10] {\footnotesize{3,z}};
        \end{tikzpicture}

        \onslide<2>
        \begin{tikzpicture}[]
            \basicbtree{show-connections}
            \node[below=0.2 of node5] {\footnotesize{1,a}};
            \node[below=1.8 of node2] {\footnotesize{...}};
            \node[below=0.2 of node6] {\footnotesize{1,z}};

            \node[below=0.2 of node7] {\footnotesize{2,a}};
            \node[below=1.8 of node3] {\footnotesize{...}};
            \node[below=0.2 of node8] {\footnotesize{2,z}};

            \node[below=0.2 of node9] {\footnotesize{3,a}};
            \node[below=1.8 of node4] {\footnotesize{...}};
            \node[below=0.2 of node10] {\footnotesize{3,z}};

            \draw[selected-path] (node1.center) -- (node2.center) -- (node5.center);
        \end{tikzpicture}

        \onslide<3>
        \begin{tikzpicture}[]
            \basicbtree{show-connections}
            \node[below=0.2 of node5] {\footnotesize{1,a}};
            \node[below=1.8 of node2] {\footnotesize{...}};
            \node[below=0.2 of node6] {\footnotesize{1,z}};

            \node[below=0.2 of node7] {\footnotesize{2,a}};
            \node[below=1.8 of node3] {\footnotesize{...}};
            \node[below=0.2 of node8] {\footnotesize{2,z}};

            \node[below=0.2 of node9] {\footnotesize{3,a}};
            \node[below=1.8 of node4] {\footnotesize{...}};
            \node[below=0.2 of node10] {\footnotesize{3,z}};

            \draw[selected-path] (node1.center) -- (node3.center) -- (node7.center);
        \end{tikzpicture}

        \onslide<4>
        \begin{tikzpicture}[]
            \basicbtree{show-connections}
            \node[below=0.2 of node5] {\footnotesize{1,a}};
            \node[below=1.8 of node2] {\footnotesize{...}};
            \node[below=0.2 of node6] {\footnotesize{1,z}};

            \node[below=0.2 of node7] {\footnotesize{2,a}};
            \node[below=1.8 of node3] {\footnotesize{...}};
            \node[below=0.2 of node8] {\footnotesize{2,z}};

            \node[below=0.2 of node9] {\footnotesize{3,a}};
            \node[below=1.8 of node4] {\footnotesize{...}};
            \node[below=0.2 of node10] {\footnotesize{3,z}};

            \draw[selected-path] (node1.center) -- (node4.center) -- (node9.center);
        \end{tikzpicture}
    \end{overprint}

    \linespread{0.5}
    \color{black}\fontsize{6pt}{0}\selectfont
        H. Leslie, R. Jain, D. Birdsall, H. Yaghmai, "Efficient search of
        multidimentionel B-trees", \\ ACM Transactions on Database Systems,
        vol. 6, no. 4, pp. 650-670, 1995
    \linespread{1.5}

    \end{center}
\end{frame}
\note{
    For not complete search keys, if a dimension is under-specified, its
    possible values must be enumerated.
}

@@

@@

        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            G. Graefe, "Sorting and indexing with partitioned B-Trees",
            Classless Inter Domain Routing, 2003
        \linespread{1.5}

@@

        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            Riegger Christian, Vincon Tobias, Petrov Ilia. (2017).
            Write-optimized indexing with partitioned b-trees. 296-300.
        \linespread{1.5}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Huanchen Zhang, David G. Andersen, Andrew Pavlo, Michael Kaminsky, Lin
        Ma, and Rui Shen. 2016. Reducing the Storage Overhead of Main-Memory
        OLTP Databases with Hybrid Indexes. In Proceedings of the 2016
        International Conference on Management of Data (SIGMOD ’16).
        Association for Computing Machinery, New York, NY, USA, 1567–1581.
    \linespread{1.5}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Levandoski Justin, Lomet David, Sengupta Sudipta. (2012). The Bw-Tree:
        A B-tree for New Hardware Platforms. Proceedings - International
        Conference on Data Engineering.
    \linespread{1.5}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Zhou Xinjing, Shou Lidan, Chen Ke, Hu Wei, Chen Gang. (2019). DPTree:
        differential indexing for persistent memory. Proceedings of the VLDB
        Endowment.
    \linespread{1.5}

@@

@@

    \begin{center}
    \textbf{Learned Index}
    \vspace{1.0cm}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Kraska Tim, Alex Beutel, Ed H. Chi, Jeffrey Dean, and Neoklis
        Polyzotis. "The case for learned index structures." In Proceedings of
        the 2018 international conference on management of data.
    \linespread{1.5}

    \end{center}
\end{frame}
\note{
}

@@

    \begin{center}
    \textbf{Learned Index}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Zhou Xinjing, Shou Lidan, Chen Ke, Hu Wei, Chen Gang. (2019). DPTree:
        differential indexing for persistent memory. Proceedings of the VLDB
        Endowment.
    \linespread{1.5}

@@

    \begin{center}
        \begin{itemize}[label={\MVRightarrow}]
            \item <+-> Buffered (Lazy) B-Tree
            \item <+-> Cache conscious B-Tree
            \item <+-> Distributed B-Tree
            \item <+-> UB-Tree
            \item <+-> Skip List
            \item <+-> LSM Tree
            \item <+-> Trie
            \item <+-> ...
        \end{itemize}
    \end{center}
\end{frame}
\note{
    Buffered B-Tree are used in WiredTiger.
    Skip list are used also in WiredTiger as in memory structure, in Apache
    Cassandra again as in memory, and in Redis.
}

>>>>

    \begin{itemize}[label={\MVRightarrow}]
        \item <+-> Comer D. (1979). The Ubiquitous B-Tree. \\
            Computing Surveys, 11 (2).
        \item <+-> Graefe G. (2011). Modern B-Tree Techniques. \\
            Foundations and Trends in Databases.
    \end{itemize}
\end{frame}

@@

    \begin{itemize}[label={\MVRightarrow}]
        \item RUM conjecture
        \item B-Tree internals
        \begin{itemize}[label={}]
            \item Key normalization
            \item Prefix/Suffix truncation
            \item SB-Tree
        \end{itemize}
        \item B-Tree as a component
        \begin{itemize}[label={}]
            \item Hybrid index
            \item Bw-Tree
            \item Partitioned B-Tree
        \end{itemize}
        \item Learned Indexes
        \item Everything else
    \end{itemize}
\end{frame}

\section{RUM conjecture}

@@

\section{B-Tree internals}

@@

    \begin{center}
    \textbf{B-Tree $\,\to\,$ B$^+$Tree}
    \vspace{1cm}

@@

            \draw[->, btree-line] ([yshift=0.1cm]sep33.east) -- ([yshift=0.1cm]sep42.west);
            \draw[->, btree-line] ([yshift=0.1cm]sep32.east) -- ([yshift=0.1cm]sep44.west);
            \draw[<-, btree-line] ([yshift=-0.1cm]sep33.east) -- ([yshift=-0.1cm]sep42.west);
            \draw[<-, btree-line] ([yshift=-0.1cm]sep32.east) -- ([yshift=-0.1cm]sep44.west);
            \coordinate[right=0.5cm of sep34.east] (link-sep34);
            \draw[->, btree-line] ([yshift=0.1cm]sep34.east) -- ([yshift=0.1cm]link-sep34.west);
            \coordinate[left=0.5cm of sep43.west] (link-sep43);
            \draw[->, btree-line] ([yshift=0.1cm]sep43.west) -- ([yshift=0.1cm]link-sep43.east);
        \end{tikzpicture}
    \end{overprint}

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Lehman P., Yao S. (1981). Efficient Locking for Concurrent Operations on B-Trees. \\
        ACM Transactions on Database Systems.
    \linespread{1.5}
    \end{center}

@@

    \begin{flushleft}
        \begin{displayquote}
            "By adding periodic rebuilding of the tree, we obtain a data
            structure that is theoreticaly superior to standard B-trees in many
            ways. Our results suggest that rebalancing on deletion no only
            unnecessary but may be harmful."
        \end{displayquote}
        \footnotesize{Sen S., Tarjan R. E. (2009). Deletion without rebalancing
            in multiway search trees ACM Transactions on Database Systems}
    \end{flushleft}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Singleton, R. C. (1969). An efficient algorithm for sorting with minimal storage. \\
        Communications of the ACM.
    \linespread{1.5}

@@

        %\onslide<5>
        %\begin{center}
        %\begin{tikzpicture}[]
            %\btreenodewithitems{1}{8}{5}{}{btree-leaf}
            %\foreach \i in {1,...,4}
            %{
                %\coordinate[below=1cm of key1\i] (start\i);
                %\draw[->, btree-line, line width=1.5pt, shorten >= 2mm] (start\i) -- (key1\i);
            %}
        %\end{tikzpicture}
        %\end{center}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Bayer, R., Unterauer, K. (1977). Prefix B-trees. \\
        ACM Transactions on Database Systems.
    \linespread{1.5}

    \end{center}
\end{frame}
\note{
    Not a normalized version, but whole column version is in PG12.
}

    \begin{overprint}[6cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \btreenodewithitems{1}{5}{2}{}{btree-branch}
            \btreenodewithitems{2}{5}{3}{below=1cm of key11, xshift=-2cm}{btree-leaf}
            \btreenodewithitems{3}{5}{3}{below=1cm of key11, xshift=2cm}{btree-leaf}
            \draw[->, btree-line] (node1.south)
                .. controls ([yshift=-1cm] node1) and ([yshift=1.2cm] node2) ..
                (node2.north);
            \draw[->, btree-line] (node1.south)
                .. controls ([yshift=-1cm] node1) and ([yshift=1.2cm] node3) ..
                (node3.north);

            \node[btree-key, below=-0.5cm of key13.center] (promoted) {};
            \draw[->, btree-line, color=red, line width=0.5mm] (key31.west)
                .. controls ([xshift=-1cm] key31) and (promoted) ..
                (promoted.center);
        \end{tikzpicture}

\end{frame}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Graefe G., Larson P. (2001). B-tree indexes and CPU caches. \\
        International Conference on Data Engineering.
    \linespread{1.5}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        O’Neil P. E. (1992). The SB-tree: An index-sequential structure for high performance sequential access. \\
        Acta Informatica.
    \linespread{1.5}

@@

    %\begin{center}
    %\textbf{Multidimentional Access Method}
    %\vspace{1cm}

    %\begin{overprint}[8cm]
        %\onslide<1>
        %\begin{tikzpicture}[]
            %\basicbtree{show-connections}
            %\node[below=0.2 of node5] {\footnotesize{1,a}};
            %\node[below=1.8 of node2] {\footnotesize{...}};
            %\node[below=0.2 of node6] {\footnotesize{1,z}};

            %\node[below=0.2 of node7] {\footnotesize{2,a}};
            %\node[below=1.8 of node3] {\footnotesize{...}};
            %\node[below=0.2 of node8] {\footnotesize{2,z}};

            %\node[below=0.2 of node9] {\footnotesize{3,a}};
            %\node[below=1.8 of node4] {\footnotesize{...}};
            %\node[below=0.2 of node10] {\footnotesize{3,z}};
        %\end{tikzpicture}

        %\onslide<2>
        %\begin{tikzpicture}[]
            %\basicbtree{show-connections}
            %\node[below=0.2 of node5] {\footnotesize{1,a}};
            %\node[below=1.8 of node2] {\footnotesize{...}};
            %\node[below=0.2 of node6] {\footnotesize{1,z}};

            %\node[below=0.2 of node7] {\footnotesize{2,a}};
            %\node[below=1.8 of node3] {\footnotesize{...}};
            %\node[below=0.2 of node8] {\footnotesize{2,z}};

            %\node[below=0.2 of node9] {\footnotesize{3,a}};
            %\node[below=1.8 of node4] {\footnotesize{...}};
            %\node[below=0.2 of node10] {\footnotesize{3,z}};

            %\draw[selected-path] (node1.center) -- (node2.center) -- (node5.center);
        %\end{tikzpicture}

        %\onslide<3>
        %\begin{tikzpicture}[]
            %\basicbtree{show-connections}
            %\node[below=0.2 of node5] {\footnotesize{1,a}};
            %\node[below=1.8 of node2] {\footnotesize{...}};
            %\node[below=0.2 of node6] {\footnotesize{1,z}};

            %\node[below=0.2 of node7] {\footnotesize{2,a}};
            %\node[below=1.8 of node3] {\footnotesize{...}};
            %\node[below=0.2 of node8] {\footnotesize{2,z}};

            %\node[below=0.2 of node9] {\footnotesize{3,a}};
            %\node[below=1.8 of node4] {\footnotesize{...}};
            %\node[below=0.2 of node10] {\footnotesize{3,z}};

            %\draw[selected-path] (node1.center) -- (node3.center) -- (node7.center);
        %\end{tikzpicture}

        %\onslide<4>
        %\begin{tikzpicture}[]
            %\basicbtree{show-connections}
            %\node[below=0.2 of node5] {\footnotesize{1,a}};
            %\node[below=1.8 of node2] {\footnotesize{...}};
            %\node[below=0.2 of node6] {\footnotesize{1,z}};

            %\node[below=0.2 of node7] {\footnotesize{2,a}};
            %\node[below=1.8 of node3] {\footnotesize{...}};
            %\node[below=0.2 of node8] {\footnotesize{2,z}};

            %\node[below=0.2 of node9] {\footnotesize{3,a}};
            %\node[below=1.8 of node4] {\footnotesize{...}};
            %\node[below=0.2 of node10] {\footnotesize{3,z}};

            %\draw[selected-path] (node1.center) -- (node4.center) -- (node9.center);
        %\end{tikzpicture}
    %\end{overprint}

    %\linespread{0.5}
    %\color{black}\fontsize{6pt}{0}\selectfont
        %Leslie, H., Jain, R., Birdsall, D., Yaghmai, H. (1995). Efficient search of multi-dimensional b-trees. VLDB.
    %\linespread{1.5}

    %\end{center}
%\end{frame}
%\note{
    %For not complete search keys, if a dimension is under-specified, its
    %possible values must be enumerated.
%}

@@

\section{B-Tree as a component}

    \def\arraystretch{1.5}
    \begin{overprint}[8cm]
        \onslide<1>
        \begin{tabular}{cccc}
            \colorbox{white}{B-Tree} & \colorbox{white}{B$^{+}$-Tree} & \colorbox{white}{B$_{link}$-Tree} & \colorbox{white}{DPTree} \\
            wB$^{+}$-Tree & NV-Tree & FPTree & FASTFAIR \\
            HiKV & Masstree & Skip List & ART \\
            WORT & CDDS-Tree & \colorbox{white}{Bw-Tree} & HOT \\
            KISS-Tree & VAST-Tree & FAST & HV-Tree \\
            UB-Tree & LHAM & \colorbox{white}{PBT} & \colorbox{white}{Hybrid B$^{+}$-Tree}
        \end{tabular}

        \onslide<2>
        \begin{tabular}{cccc}
            \colorbox{red!20}{B-Tree} & \colorbox{red!20}{B$^{+}$-Tree} & \colorbox{red!20}{B$_{link}$-Tree} & DPTree \\
            wB$^{+}$-Tree & NV-Tree & FPTree & FASTFAIR \\
            HiKV & Masstree & Skip List & ART \\
            WORT & CDDS-Tree & \colorbox{white}{Bw-Tree} & HOT \\
            KISS-Tree & VAST-Tree & FAST & HV-Tree \\
            UB-Tree & LHAM & \colorbox{white}{PBT} & \colorbox{white}{Hybrid B$^{+}$-Tree}
        \end{tabular}

        \onslide<3>
        \begin{tabular}{cccc}
            \colorbox{red!20}{B-Tree} & \colorbox{red!20}{B$^{+}$-Tree} & \colorbox{red!20}{B$_{link}$-Tree} & \colorbox{red!20}{DPTree} \\
            wB$^{+}$-Tree & NV-Tree & FPTree & FASTFAIR \\
            HiKV & Masstree & Skip List & ART \\
            WORT & CDDS-Tree & \colorbox{red!20}{Bw-Tree} & HOT \\
            KISS-Tree & VAST-Tree & FAST & HV-Tree \\
            UB-Tree & LHAM & \colorbox{red!20}{PBT} & \colorbox{red!20}{Hybrid B$^{+}$-Tree}
        \end{tabular}

\end{frame}

@@

        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            Graefe G. (2003). Sorting and indexing with partitioned B-Trees. \\
            Classless Inter Domain Routing
        \linespread{1.5}

@@

        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            Riegger C., Vincon T., Petrov I. (2017). Write-optimized indexing with partitioned b-trees. \\
            Proceedings of the 19th International Conference on Information Integration and Web-based Applications \& Services
        \linespread{1.5}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Zhang H., Andersen D. G., Pavlo A., Kaminsky M., Ma L., Shen R. (2016). \\
        Reducing the Storage Overhead of Main-Memory OLTP Databases with Hybrid
        Indexes. \\
        Proceedings of the 2016 International Conference on Management of Data
        (SIGMOD ’16)
    \linespread{1.5}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Levandoski J., Lomet D., Sengupta S. (2012). The Bw-Tree: A B-tree for
        New Hardware Platforms. \\
        International Conference on Data Engineering.
    \linespread{1.5}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
         Zhou Xinjing, Shou Lidan, Chen Ke, Hu Wei, Chen Gang. (2019). DPTree:
         differential indexing for persistent memory. \\
         Proceedings of the VLDB Endowment.
    \linespread{1.5}

@@

\section{Learned Indexes}

@@

    \begin{center}
    \textbf{Learned Indexes}
    \vspace{1.0cm}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Kraska T., Beutel A., Chi E. H., Dean J., Polyzotis N. (2018). The Case
        for Learned Index Structures. \\
        2018 International Conference on Management of Data (SIGMOD ‘18).
    \linespread{1.5}

@@

    \begin{center}
    \textbf{Data approximation}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Galakatos, A., Markovitch, M., Binnig, C., Fonseca, R., Kraska, T.
        (2019). Fiting-tree: A data-aware index structure. \\
        Proceedings of the 2019 International Conference on Management of Data.
    \linespread{1.5}

@@

    \begin{center}
        \begin{itemize}[label={\MVRightarrow}]
            \item <+-> Buffered (Lazy) B-Tree
            \item <+-> Cache conscious B-Tree
            \item <+-> Distributed B-Tree
            \item <+-> UB-Tree
            \item <+-> MDAM
            \item <+-> LSM Tree
            \item <+-> Trie
            \item <+-> ...
        \end{itemize}
    \end{center}
\end{frame}
\note{
    Buffered B-Tree are used in WiredTiger.
    Skip list are used also in WiredTiger as in memory structure, in Apache
    Cassandra again as in memory, and in Redis.
}
